---
title: "Zooplankton salinity relationships in Suisun"
author: "Sam Bashevkin, updated by Rosemary hartman"
date: "8/26/2022 - updated spring 2025"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    )
    })
    rmarkdown::render(
      input,
      envir = globalenv()
      output_dir = here::here('docs'),
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

# Load packages

```{r, message=FALSE, warning=FALSE}
library(conflicted)
library(MASS)
library(dplyr)
library(zooper)
library(lubridate)
library(readr)
library(tidyr)
library(ggplot2)
library(sf)
library(readxl)
library(stringr)
library(mgcv)
library(purrr)
library(deltamapr)
library(scales)
library(here)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```



# Load and wrangle data

We used all the available zooplankton data in the 'zooper' integrated dataset from EMP, Summer Twonet, 20-mm and Fall Midwater Trawl. We ended up restricting it to 1995-2022.

Available here: https://github.com/InteragencyEcologicalProgram/zooper

```{r}
zoop_data<-Zoopsynther(Data_type="Community", Sources=c("EMP", "STN", "20mm", "FMWT"), Time_consistency = FALSE)
```


Read in zoop mass conversions. These are conversions based on an average weight per taxon and life stage. Some are better than others, particularly "other cladocera" which can have a huge range of sizes.

The zoop_groups file is a crosswalk that lines up the names of each taxa to the groups used by the bioenergetic model.

```{r}


#Mass conversions
zoop_mass_conversions<-read_excel(here("Data/Biomass conversions.xlsx"), sheet="Micro and Meso-zooplankton")%>%
  mutate(Taxname=case_when(Taxname=="Sinocalanus"~"Sinocalanus doerrii", # Change to help this match to zoop data
                           TRUE ~ Taxname),
         Taxlifestage=paste(Taxname, Lifestage))%>%
  select(Taxlifestage, CarbonWeight_ug)

#Read in zoop groupings
zoop_groups<-read_csv(here("Data/zoopcrosswalk2.csv"), col_types=cols_only(Taxlifestage="c", IBMR="c"))%>%
  distinct()

```


Start processing the zoop data
```{r}
zoop_data_mass<-zoop_data%>%
  mutate(Taxlifestage=str_remove(Taxlifestage, fixed("_UnID")), doy = yday(Date), Month = month(Date))%>%
  filter(!Undersampled, SizeClass != "Macro", Year >1995, Month %in% c(6:10)) %>% #only meso and microzooplankton, not undersampled categories. Also just summer and fall.
  
  mutate(Taxlifestage=recode(Taxlifestage, `Synchaeta bicornis Adult`="Synchaeta Adult", # Change some names to match to biomass conversion dataset
                             `Pseudodiaptomus Adult`="Pseudodiaptomus forbesi Adult",
                             `Acanthocyclops vernalis Adult`="Acanthocyclops Adult"))%>%
  
  left_join(zoop_mass_conversions, by="Taxlifestage")%>% # Add biomass conversions
  left_join(zoop_groups, by="Taxlifestage")%>% # Add IBMR categories
  mutate(BPUE=CPUE*CarbonWeight_ug)%>% # Convert CPUE to biomass per unit effort
  
  filter(!is.na(BPUE) & !is.na(Latitude) & !is.na(Longitude) & !is.na(SalSurf))%>% # Removes any data without BPUE, which is currently restricted to Rotifera Adult, Copepoda Larva, and H. longirostris from STN. Also removes 20mm and EMP EZ stations without coordinates
  
  group_by(IBMR)%>%
  select(SampleID, Station, Latitude, Longitude, SalSurf, Date, Year, IBMR, BPUE)%>%
  group_by(across(-BPUE))%>%
  summarise(BPUE=sum(BPUE), .groups="drop")%>% # Sum each IBMR categories
  
  st_as_sf(coords=c("Longitude", "Latitude"), crs=4326)%>% #convert to spatial data frame
  st_transform(crs=st_crs(deltamapr::R_DSIBM)) %>%  #add IBMR regions
  st_join(deltamapr::R_DSIBM %>%
            select(SUBREGION)) %>%
  st_drop_geometry() %>% 
  filter(SUBREGION %in% c("NW Suisun","SW Suisun","NE Suisun","SE Suisun","Confluence", "Suisun Marsh", "Lower Sacramento River"))%>% #just subregions of interest
  mutate(doy=yday(Date), #Day of year
         Month=month(Date), # Month
         Year_fac=factor(Year), # Factor year for model random effect
         Station_fac=factor(Station), # Factor station for model random effect
         across(c(SalSurf, doy), list(s=~(.x-mean(.x))/sd(.x))), # Center and standardize predictors
         BPUE_log1p=log(BPUE+1)) # log1p transform BPUE for model

```

Check sample size
```{r message=FALSE, warning=FALSE}
zoop_sample_size <- zoop_data_mass %>% 
  group_by(SampleID,Year,Month,SUBREGION,Station) %>% 
  summarise(BPUE=sum(BPUE)) %>% 
  mutate(Samplesize=1) %>%
  group_by(Year, Month, SUBREGION) %>% 
  summarise(mean_BPUE=mean(BPUE),Samplesize=sum(Samplesize)) %>%
  filter(Year>=1995)

ggplot(zoop_sample_size, aes(x=Year, y=Month, fill=Samplesize))+
  geom_tile()+
  scale_y_continuous(breaks=1:12, labels=month(1:12, label=T))+
  scale_fill_viridis_c(breaks=c(1,5,10,15,20))+
  facet_wrap(~SUBREGION)+
  theme_bw()
```

All the remaining brackish regions have sufficient sample size with the exception of NE Suisun. As such, NE Suisun is to be combined with SE Suisun while the rest of the regions are to be analyzed on their own.

Create a new column with IBMR edited regions to accomodate combination of NE and SE Suisun regions.
```{r}
zoop_data_mass$Subregion_edit<-ifelse(zoop_data_mass$SUBREGION%in%c("NE Suisun", "SE Suisun"), "East Suisun", zoop_data_mass$SUBREGION)

save(zoop_data_mass, file = "data/zoop_data_mass.RData")

```

A few graphs to compare actual zooplankton data to model outputs later.

```{r}
zoopmass2016 = filter(zoop_data_mass, Year == 2016) %>%
  filter(Month %in% c(6:10)) %>%
  group_by(Subregion_edit, IBMR) %>%
  summarise(Biomass = exp(mean(BPUE_log1p)))

ggplot(zoopmass2016, aes(x = Subregion_edit,y = Biomass, fill = IBMR))+ geom_col()

zoopmass20162 = mutate(zoopmass2016, 
                       BigRegion = case_when(Subregion_edit %in% c("East Suisun", "NW Suisun", "SW Suisun") ~ "Suisun Bay", 
                                             Subregion_edit == "Suisun Marsh" ~ "Suisun Marsh",
                                             Subregion_edit == "Confluence" ~ "Confluence")) %>%
  filter(!is.na(BigRegion)) %>%
  group_by(BigRegion, IBMR) %>%
  summarize(Biomass = mean(Biomass))

pal = c("red", "darkred", "green",  "darkblue", "cyan", "pink", "darkgreen", "purple", "tan", "orange", "palegreen", "salmon", "grey")
ggplot(zoopmass20162, aes(x = BigRegion,y = Biomass, fill = IBMR))+ geom_col()+
  scale_fill_manual(values = pal)
ggplot(filter(zoopmass20162, IBMR %in% c("pdiapfor", "pdiapjuv")), aes(x = BigRegion, y = Biomass, fill = IBMR))+
  geom_col()

ggplot(filter(zoop_data_mass, Subregion_edit == "Confluence", IBMR == "pdiapfor"),
       aes(x = SalSurf, y = BPUE_log1p))+ geom_point()+ geom_smooth()

ggplot(filter(zoop_data_mass, Subregion_edit %in% c("Confluence", "Suisun Marsh"), IBMR == "pdiapfor",
  Month %in% c(6:10)),
       aes(x = Subregion_edit, y = BPUE))+ geom_boxplot()

summary(filter(zoop_data_mass, Subregion_edit == "Confluence", IBMR == "pdiapfor")$BPUE)

```



# Model

## Prediction data

Set up prediction data for model
```{r}
# Min year to start models
year_min<-1995

#This sets up the data for predicting later. It's all combintations of IBMR, salinity, and region

newdata_function<-function(region, data=zoop_data_mass, quant=0.99){
  
  lower<-(1-quant)/(2)
  upper<-1-lower
  
  data_filt<-data%>%
    filter(Subregion_edit%in%region & Year >= year_min)
  
  # Calculate monthly quantiles of salinity
  month_sal<-data_filt%>%
    group_by(Month)%>%
    summarise(l=quantile(SalSurf, lower),
              u=quantile(SalSurf, upper), .groups="drop")
  
  newdata<-expand_grid(date=seq(ymd("2001-06-01"), ymd("2001-10-31"), by = "day"), # all the days in a non-leap year
                       SalSurf=seq(round(min(data_filt$SalSurf), 1), 
                                   round(max(data_filt$SalSurf), 1), by=0.1))%>% # Salinity sequence nicely rounded to 1 decimal
    mutate(Month=month(date),
           doy=yday(date), # Day of year
           SalSurf_s=(SalSurf-mean(data$SalSurf))/sd(data$SalSurf), # center and standardize salinity to match data
           doy_s=(doy-mean(data$doy))/sd(data$doy))%>% # center and standardize doy to match data
    left_join(month_sal, by="Month")%>%
    filter(SalSurf >= l & SalSurf <= u)%>% # Remove any salinity values outside the quantiles for each month
    select(Month, doy, doy_s, SalSurf, SalSurf_s)
  
}

#create data for each subregion
newdata<-map(set_names(unique(zoop_data_mass$Subregion_edit)), newdata_function)
```

## Posterior prediction function

```{r}
# Function to generate posterior predictions from a gam model
# From https://stats.stackexchange.com/questions/190348/can-i-use-bootstrapping-to-estimate-the-uncertainty-in-a-maximum-value-of-a-gam
predict_posterior<-function(model, newdata, exclude, n=499, seed=999){
  Xp <- predict(model, newdata=newdata, type="lpmatrix", exclude=exclude, newdata.guaranteed=TRUE) ## map coefs to fitted curves
  beta <- coef(model)
  Vb   <- vcov(model) ## posterior mean and cov of coefs
  set.seed(seed)
  mrand <- mvrnorm(n, beta, Vb) ## simulate n rep coef vectors from posterior
  pred<-matrix(nrow=nrow(newdata), ncol=n)
  ilink <- family(model)$linkinv
  for (i in seq_len(n)) { 
    pred[,i]   <- ilink(Xp %*% mrand[i, ])
  }
  colnames(pred)<-paste("draw", 1:n, sep="_")
  pred<-as_tibble(pred)
  return(pred)
}
```

## Model fitting

Function to run a generalized addadive model over every IBMR group and region.

Each model is the log-transformed BPUE as predicted by the tensor product smooth of salinity and day of year, with year (as a factor), and station as random effects.

```{r}
sal_model<-function(group,region,new_data=newdata){
  
  cat("<<<<<<<<<<<<<<<<<<<<<<< modeling", group, region, ">>>>>>>>>>>>>>>>>>>>>>>>>\n\n")
  
  new_data<-new_data[[region]]
  
  data<-filter(zoop_data_mass, IBMR==group & Subregion_edit==region & Year>=year_min)
  
  par(mfrow=c(2,2))
  
  if(length(unique(data$Station_fac))>1){
    model<-gam(BPUE_log1p ~ te(SalSurf_s, doy_s, k=c(5,5), bs=c("cs", "cs")) + 
                 s(Year_fac, bs="re") + s(Station_fac, bs="re"),
               data=data, 
               method="REML")
    
    random_effects<-c("s(Year_fac)", "s(Station_fac)")
    
  }else{
    
    model<-gam(BPUE_log1p ~ te(SalSurf_s, doy_s, k=c(5,5), bs=c("cs", "cs")) + 
                 s(Year_fac, bs="re"),
               data=data, 
               method="REML")
    
    random_effects<-c("s(Year_fac)")
  }
  
  cat("-------------gam check-------------\n")
 # gam.check(model)
  
  cat("\n\n-------------summary-------------\n")
  print(summary(model))
  
  sal<-predict_posterior(model, new_data, random_effects)%>%
    bind_cols(new_data%>% # Add covariate columns before these columns
                select(-doy_s, -SalSurf_s), 
              .)
  return(sal)
}
```

Apply model to all groups and regions

```{r}

#Set up data frame of all regions and IBMR categories
#I had issues with the computer crashing when I used the lower sac river region with Limnothona, so I took it out.
model_factors<-expand_grid(IBMR=unique(zoop_data_mass$IBMR),
                           Subregion_edit=unique(zoop_data_mass$Subregion_edit))%>%
  mutate(IBMR=set_names(IBMR, paste(IBMR, Subregion_edit))) %>%
  filter(Subregion_edit != "Lower Sacramento River")

#now run the model over every region and IBMR group. The output is a list of what biomass
#you get for each salinity that is normally seen in each region.
sal_conversionsa<-pmap_dfr(model_factors, function(IBMR, Subregion_edit) sal_model(IBMR, Subregion_edit), .id = "IBMR_region")

#turn it into a data fame and make it more useful. 
sal_conversions = bind_rows(sal_conversionsa)%>%
  mutate(IBMR=sapply(IBMR_region, function(x) str_split(x, " ", n=2)[[1]][1]),
         Region=factor(sapply(IBMR_region, function(x) str_split(x, " ", n=2)[[1]][2]),
                       levels=c("Confluence", "Suisun Marsh", "East Suisun", 
                                "NW Suisun", "SW Suisun", "Lower Sacramento River")),
         Month=as.integer(Month))%>%
  select(-IBMR_region)%>%
  relocate(Region, Month, IBMR, SalSurf)

#save the results
#save(sal_conversions,  file = "outputs/salinityConversions.RData")
```


# Apply model

Now we use the table of predictions of biomass at each salnitiy and the modeled salintiy to see what biomass would be in each mdoel scenario.

Load in SMSCG modeled salinity
```{r, message=FALSE}
#OK, here is where we need to upload new salinities. 

#The salnity files are in a bit of a weird format, so it took some doing to straeighen them out.
extract_after_first_space <- function(input_string) {
  return(gsub("^\\S+\\s+", "", input_string))
}

salinity2016names = read_csv("data/2016_combined_salinity_ibmr.csv", col_names = F)[c(1,2),]

sal2016names = t(salinity2016names) %>%
  as.data.frame() %>%
  mutate(Name = paste(V1, V2))

salinity2016data = read_csv("data/2016_combined_salinity_ibmr.csv", col_names = F, skip =3)

names(salinity2016data) = sal2016names$Name

sal2016 = pivot_longer(salinity2016data, cols = c(`10800-3-m Confluence`:last_col()), names_to = "Scenario2", values_to = "Salinity") %>%
  mutate(Region = extract_after_first_space(Scenario2)) %>%
  rename(Date = `scenario region`)

sal2016$Scenariox = str_split(sal2016$Scenario2, " ")
sal2016$Scenario = sapply(sal2016$Scenariox, first)

#sal2016 is the good version. 


```

OK, now do it for 2024

```{r}

salinity2024names = read_csv("data/2024_combined_salinity_ibmr.csv", col_names = F)[c(1,2),]

sal2024names = t(salinity2016names) %>%
  as.data.frame() %>%
  mutate(Name = paste(V1, V2))

salinity2024data = read_csv("data/2024_combined_salinity_ibmr.csv", col_names = F, skip =3)

names(salinity2024data) = sal2024names$Name

sal2024 = pivot_longer(salinity2024data, cols = c(`10800-3-m Confluence`:last_col()), names_to = "Scenario2", values_to = "Salinity") %>%
  mutate(Region = extract_after_first_space(Scenario2)) %>%
  rename(Date = `scenario region`)

sal2024$Scenariox = str_split(sal2024$Scenario2, " ")
sal2024$Scenario = sapply(sal2024$Scenariox, first)

```



Calculate zoop abundance difference between each scenario and the baseline.

This was origionally one very long process, but I broke it into four steps because it was too much all at once and I kept having issues.


```{r}

#first take the salinity data and expand it to include all IBMR groups.
zoop_saladjusted2016<-sal2016%>%
  select(-Scenario2, -Scenariox) %>%
  distinct() %>%
  mutate(Subregion_edit=if_else(Region%in%c("NE Suisun", "SE Suisun"), "East Suisun", as.character(Region))) %>%
mutate(Salinity=as.character(Salinity), Year = 2016,  DOY = yday(Date))%>%
  left_join(distinct(model_factors)) %>%
  ungroup()

#join in salinity conversions so you get biomass
step2= zoop_saladjusted2016%>%
  left_join(sal_conversions%>%
              mutate(SalSurf=as.character(SalSurf)),
            by=c("Subregion_edit"="Region",
                 "DOY"="doy",
                 "Salinity"="SalSurf",
                 "IBMR"="IBMR"))


#back-transform biomass from the log version to the real version
step3 = step2%>%
  mutate(across(starts_with("draw_"), ~exp(.x)-1))%>%
  pivot_longer(starts_with("draw_"), names_prefix="draw_", names_to="draw", values_to="fit")%>%
  mutate(fit=if_else(fit<0, 0, fit))%>%
  pivot_wider(names_from="Scenario", values_from="fit")

#calculate the median and 95% confidence interval across simulations.
step4 = step3%>%
  mutate(across(c(`10800-3-m`:last_col()), ~.x/as.numeric(Salinity)))%>%
  group_by(Region, Year, DOY, IBMR)%>%
  summarise(across(c(`10800-3-m`:last_col()), 
                   list(median=~median(.x, na.rm=T), 
                        l95=~quantile(.x, 0.025, na.rm=T), 
                        u95=~quantile(.x, 0.975, na.rm=T))), 
            .groups="drop")


write_csv(step4, here("Outputs", "sdm zoop sal adjustments.csv"))
step4 = read_csv("Outputs/sdm zoop sal adjustments.csv")
```


This next step is just a tedious process of putting it in a better format for graphing.

```{r}

zoopsmedian = pivot_longer(step4,  cols = c(`10800-3-m_median`:last_col()),
       names_to = "Scenario", values_to = "Biomass") %>%
filter(!str_detect(Scenario, "95"), Region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_median"))



zoopsU95 = pivot_longer(step4,  cols = c(`10800-3-m_median`:last_col()),
       names_to = "Scenario", values_to = "BiomassU95") %>%
filter(str_detect(Scenario, "u95"), Region %in% c("Confluence", "NW Suisun",  "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_u95"))


zoopsl95 = pivot_longer(step4,  cols = c(`10800-3-m_median`:last_col()),
       names_to = "Scenario", values_to = "BiomassL95") %>%
filter(str_detect(Scenario, "l95"), Region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_l95"))

zoopsconverted = left_join(zoopsmedian, zoopsU95)%>%
                             left_join(zoopsl95)

ggplot(zoopsmedian, aes(x = DOY, y = Biomass, color = Scenario)) +
  facet_grid(Region~IBMR)+ geom_line()+ scale_y_log10()


ggplot(zoopsmedian, aes(x = DOY, y = Biomass, fill = IBMR)) +
  facet_grid(Region~Scenario, scales = "free_y")+ geom_area()

```

Average biomass of pseudodiaptomus juveniles and adults in Suisun Marhs and Suisun bay.

I used the geometric mean (mean of the log-transformed pseudodiaptomus) since that levels out the super high points, but I'm not 100% that it's the rigth way to go. 

```{r}

#calculate the mean biomass
zoopsave = zoopsconverted %>%
  filter(IBMR %in% c("pdiapfor", "pdiapjuv")) %>%
  group_by(Scenario, Year, DOY, Region) %>%
  summarize(Biomass = sum(Biomass), U95 = sum(BiomassU95), L95 = sum(BiomassL95)) %>%
  group_by(Scenario,  Region) %>%
  summarize(MeanBiomass = mean(Biomass, na.rm =T), sdbiomass = sd(Biomass, na.rm=T), 
            MeanU95 = exp(mean(log(U95+0.01), na.rm =T)),
            MeanL95 = exp(mean(log(L95+0.01), na.rm =T)),
            GeoMeanBiomass = exp(mean(log(Biomass+0.01), na.rm =T)), sdgio = sd(log(Biomass+0.01), na.rm=T), lower = exp(mean(log(Biomass+0.01), na.rm =T)-sdgio),
            upper = exp(mean(log(Biomass+0.01), na.rm =T)+sdgio)) %>%
  mutate(Scenario = factor(Scenario, levels = c("Noaction", "fallx2", "10800-3-m", "10800-60d", "10800-b",
                                         "15000-3-m", "15000-60d", "15000-b", "21000")))


ggplot(zoopsave, aes(x = Scenario, y = MeanBiomass, fill = Scenario))+ geom_col()+
  facet_wrap(~Region)+scale_y_log10()

ggplot(zoopsave, aes(x = Scenario, y = GeoMeanBiomass, fill = Scenario))+ geom_col()+
  geom_errorbar(aes(ymin = MeanL95, ymax = MeanU95))+
  facet_wrap(~Region)+scale_y_log10()+
  theme(axis.text.x = element_text(angle = 90))

summary(zoopsmedian$Biomass)

```

Repeat it for the 2024 data


```{r}
zoop_saladjusted2024<-sal2024%>%
  select(-Scenario2, -Scenariox) %>%
  distinct() %>%
  mutate(Subregion_edit=if_else(Region%in%c("NE Suisun", "SE Suisun"), "East Suisun", as.character(Region))) %>%
mutate(Salinity=as.character(Salinity), Year = 2024,  DOY = yday(Date))%>%
  left_join(distinct(model_factors)) %>%
  ungroup()%>%
  filter(Subregion_edit %in% c("Confluence", "East Suisun", "NW Suisun", "SW Suisun", "Suisun Marsh"),
         DOY %in% c(152:304))

step2_2024= zoop_saladjusted2024 %>%
  left_join(sal_conversions%>%
              mutate(SalSurf=as.character(SalSurf)),
            by=c("Subregion_edit"="Region",
                 "DOY"="doy",
                 "Salinity"="SalSurf",
                 "IBMR"="IBMR"))



step3_2024 = step2_2024%>%
  mutate(across(starts_with("draw_"), ~exp(.x)-1))%>%
  pivot_longer(starts_with("draw_"), names_prefix="draw_", names_to="draw", values_to="fit")%>%
  mutate(fit=if_else(fit<0, 0, fit))%>%
  pivot_wider(names_from="Scenario", values_from="fit")

step4_2024 = step3_2024%>%
  mutate(across(c(`10800-3-m`:last_col()), ~.x/as.numeric(Salinity)))%>%
  group_by(Region, Year, DOY, IBMR)%>%
  summarise(across(c(`10800-3-m`:last_col()), 
                   list(median=~median(.x, na.rm=T), 
                        l95=~quantile(.x, 0.025, na.rm=T), 
                        u95=~quantile(.x, 0.975, na.rm=T))), 
            .groups="drop")

write_csv(step4_2024, here("Outputs", "sdm zoop sal adjustments2024.csv"))
step4_2024 = read_csv("Outputs/sdm zoop sal adjustments2024.csv")


save(sal_conversions, sal2016, step4, step4_2024, sal2024, file = "outputs/SalinityStuff.RData")

#write.csv(sal_conversions, "sal_conversions.csv")

```



```{r}
zoopsmedian2024 = pivot_longer(step4_2024,  cols = c(`10800-3-m_median`:last_col()),
       names_to = "Scenario", values_to = "Biomass") %>%
filter(!str_detect(Scenario, "95"), Region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_median"))



zoopsU952024 = pivot_longer(step4_2024,  cols = c(`10800-3-m_median`:last_col()),
       names_to = "Scenario", values_to = "BiomassU95") %>%
filter(str_detect(Scenario, "u95"), Region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_u95"))


zoopsl952024 = pivot_longer(step4_2024,  cols = c(`10800-3-m_median`:last_col()),
       names_to = "Scenario", values_to = "BiomassL95") %>%
filter(str_detect(Scenario, "l95"), Region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_l95"))

zoopsconverted2024 = left_join(zoopsmedian2024, zoopsU952024)%>%
                             left_join(zoopsl952024)

zoopsave2024 = zoopsconverted2024 %>%
  filter(IBMR %in% c("pdiapfor", "pdiapjuv")) %>%
  group_by(Scenario, Year, DOY, Region) %>%
  summarize(Biomass = sum(Biomass), U95 = sum(BiomassU95), L95 = sum(BiomassL95)) %>%
  group_by(Scenario,  Region) %>%
  summarize(MeanBiomass = mean(Biomass, na.rm =T), sdbiomass = sd(Biomass, na.rm=T), 
            MeanU95 = exp(mean(log(U95+0.01), na.rm =T)),
            MeanL95 = exp(mean(log(L95+0.01), na.rm =T)),
            GeoMeanBiomass = exp(mean(log(Biomass+0.01), na.rm =T)), sdgio = sd(log(Biomass+0.01), na.rm=T), lower = exp(mean(log(Biomass+0.01), na.rm =T)-sdgio),
            upper = exp(mean(log(Biomass+0.01), na.rm =T)+sdgio)) %>%
  mutate(Scenario = factor(Scenario, levels = c("Noaction", "fallx2", "10800-3-m", "10800-60d", "10800-b",
                                         "15000-3-m", "15000-60d", "15000-b", "21000")))

ggplot(zoopsave2024, aes(x = Scenario, y = MeanBiomass, fill = Scenario))+ geom_col()+
  facet_wrap(~Region)+scale_y_log10()
```


Plots of all pseudodiaptomus - both years

```{r}
allzoopsave = bind_rows(mutate(zoopsave, Year = "2016"), mutate(zoopsave2024, Year = "2024"))

ggplot(allzoopsave, aes(x = Scenario, y = GeoMeanBiomass, fill = Scenario, group = Year))+ geom_col(position = "dodge", color = "grey", aes(alpha = Year))+
  scale_alpha_manual(values = c(0.6, 1))+
  facet_wrap(~Region)+scale_y_log10()+
  geom_errorbar(aes(ymin = MeanL95, ymax = MeanU95), position = "dodge")+
  theme(axis.text.x = element_text(angle = 90))+
  ylab("Mean summer Pseudodiaptomus biomass")

ggsave("plots/pseudobiomass.png", device = "png", width =10, height =7)

write.csv(allzoopsave, "outputs/allzoopsave.csv")

pseudoSuisun = allzoopsave %>%
  mutate(BigRegion = case_when(Region %in% c("SE Suisun", "NW Suisun", "NE Suisun", "SW Suisun") ~ "Suisun Bay",
                               Region == "Suisun Marsh" ~ "Suisun Marsh",
                               TRUE ~ Region)) %>%
  group_by(BigRegion, Scenario, Year) %>%
  summarize(MeanBiomass = mean(GeoMeanBiomass), MeanU95 = mean(MeanU95), MeanL95 = mean(MeanL95))

write.csv(pseudoSuisun, "outputs/pseudoSuisun.csv")


ggplot(pseudoSuisun, aes(x = Scenario, y = MeanBiomass, fill = Scenario, group = Year))+ geom_col(position = "dodge", color = "grey", aes(alpha = Year))+
  scale_alpha_manual(values = c(0.6, 1))+
  facet_wrap(~BigRegion)+scale_y_log10()+
  geom_errorbar(aes(ymin = MeanL95, ymax = MeanU95), position = "dodge")+
  theme(axis.text.x = element_text(angle = 90))+
  ylab("Mean summer Pseudodiaptomus biomass")


```


plost of all zoops all together



```{r}
zoopsmedALL = bind_rows(zoopsmedian, zoopsmedian2024) %>%
  mutate(Scenario = str_remove(Scenario, "_median")) %>%
  filter(DOY %in% c(153:305))%>%
  mutate(BigRegion = case_when(Region %in% c("SE Suisun", "NW Suisun", "NE Suisun", "SW Suisun") ~ "Suisun Bay",
                               Region == "Suisun Marsh" ~ "Suisun Marsh",
                               TRUE ~ Region)) %>%
  group_by(BigRegion, Scenario, Year, IBMR) %>%
  summarize(MeanBiomass = mean(Biomass, na.rm =T), sdbiomass = sd(Biomass, na.rm=T), 
            GeoMeanBiomass = exp(mean(log(Biomass+0.01), na.rm =T))) %>%
  mutate(Scenario = factor(Scenario, levels = c("Noaction",  "fallx2" ,"10800-3-m", "10800-60d", "10800-b",   "15000-3-m", "15000-60d", "15000-b",   "21000" )))


ggplot(zoopsmedALL, aes(x = Scenario, y = GeoMeanBiomass, fill = IBMR))+
  geom_col(color = "grey30")+
  facet_wrap(Year~BigRegion)+
  ylab("Mean summer-fall biomass (mg/m3)")+
  theme(axis.text.x = element_text(angle = 90))


x2issues = bind_rows(zoopsconverted, zoopsconverted2024) %>%
  filter(Scenario %in% c("fallx2", "Noaction"))

ggplot(filter(x2issues, Year == "2024"), aes(x = DOY, y = Biomass, fill = IBMR))+
  facet_grid(Region ~ Scenario, scales = "free_y")+
  geom_area()+
  ylab("Mean summer-fall biomass (mg/m3)")

```

What are the relationships between zooplankton and fall flow? All taxa

```{r}
library(glmmTMB)
load("data/zoop_data_mass.RData")
load("C:/Users/rhartman/OneDrive - California Department of Water Resources/salinity control gates/SFHA_synthesis/data/dayflow_w2024.RData")
zoop_mass_fall = zoop_data_mass %>%
  filter(Month %in% c(9:11)) %>%
  left_join(select(Dayflow, Date, OUT, X2)) %>%
  filter(SalSurf <6 & SalSurf >0.5)

ggplot(zoop_mass_fall, aes(x = log(OUT), y = log(BPUE)))+
  facet_wrap(~IBMR)+ geom_point()+ geom_smooth(method = "lm")

ggplot(zoop_mass_fall, aes(x = X2, y = log(BPUE)))+
  facet_wrap(~IBMR)+ geom_point()+ geom_smooth(method = "lm")

modout = function(data, taxa) {
  data = filter(data, IBMR == taxa)
  mod = glmmTMB(BPUE ~ X2  + (1|Month) + (1|Year), data = data, family = "nbinom2")
  modsum = as.data.frame(summary(mod)$coefficients$cond)
  modsum = mutate(modsum, IBMR = taxa)
  return(modsum)
}

zoopresults = data.frame(NA)
taxa = unique(zoop_data_mass$IBMR)
for(i in c(1:13))  {
  dat = modout(data = zoop_mass_fall, taxa = taxa[i])
  zoopresults = bind_rows(zoopresults, dat)
}

```

OK, now the adjusted biomass for the 2024 modeling

```{r}

hist = read_csv("data/ave_salinity_ts_hist_salinity.csv") %>%
  mutate(Scenario = "Gates+X2")

firstbit = filter(hist, time < ymd_hms("2024-06-28 16:30:00 UTC"))

NoAction = read_csv("data/ave_salinity_ts_no_op_no_x2_salinity.csv")%>%
  bind_rows(firstbit) %>%
  mutate(Scenario = "NoAction")

X2only = read_csv("data/ave_salinity_ts_no_op_x2_salinity.csv")%>%
    bind_rows(firstbit) %>%
  mutate(Scenario = "X2Only")

Gatesonly = read_csv("data/ave_salinity_ts_op_no_x2_salinity.csv")%>%
    bind_rows(firstbit) %>%
  mutate(Scenario = "GatesOnly")


dat2020 = bind_rows(hist, NoAction, X2only, Gatesonly)


#first take the salinity data and expand it to include all IBMR groups.
zoop_sal2024<-dat2020%>%
  distinct() %>%
  mutate(Subregion_edit=if_else(region%in%c("NE Suisun", "SE Suisun"), "East Suisun", as.character(region))) %>%
mutate(Salinity=as.character(mean), Year = 2016,  DOY = yday(time))%>%
  left_join(distinct(model_factors)) %>%
  ungroup()

#join in salinity conversions so you get biomass
step2= zoop_sal2024%>%
  left_join(sal_conversions%>%
              mutate(SalSurf=as.character(SalSurf)),
            by=c("Subregion_edit"="Region",
                 "DOY"="doy",
                 "Salinity"="SalSurf",
                 "IBMR"="IBMR"))


#back-transform biomass from the log version to the real version
step3 = step2%>%
  mutate(across(starts_with("draw_"), ~exp(.x)-1))%>%
  pivot_longer(starts_with("draw_"), names_prefix="draw_", names_to="draw", values_to="fit")%>%
  mutate(fit=if_else(fit<0, 0, fit))%>%
  pivot_wider(names_from="Scenario", values_from="fit")

#calculate the median and 95% confidence interval across simulations.
step424 = step3%>%
  mutate(across(c(`Gates+X2`:last_col()), ~.x/as.numeric(Salinity)))%>%
  group_by(region, Year, DOY, IBMR)%>%
  summarise(across(c(`Gates+X2`:last_col()), 
                   list(median=~median(.x, na.rm=T), 
                        l95=~quantile(.x, 0.025, na.rm=T), 
                        u95=~quantile(.x, 0.975, na.rm=T))), 
            .groups="drop")


write.csv(step424, here("Outputs", "zoop sal adjustments_2024gates.csv"), row.names = F)

zoopsmedian2024b = pivot_longer(step424,  cols = c(`Gates+X2_median`:last_col()),
       names_to = "Scenario", values_to = "Biomass") %>%
filter(!str_detect(Scenario, "95"), region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_median"))



zoopsU952024b = pivot_longer(step424,  cols = c(`Gates+X2_median`:last_col()),
       names_to = "Scenario", values_to = "BiomassU95") %>%
filter(str_detect(Scenario, "u95"), region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_u95"))


zoopsl952024b = pivot_longer(step424,  cols = c(`Gates+X2_median`:last_col()),
       names_to = "Scenario", values_to = "BiomassL95") %>%
filter(str_detect(Scenario, "l95"), region %in% c("Confluence", "NW Suisun", "SW Suisun",
                                                  "NE Suisun", "East Suisun", "Suisun Marsh",
                                                  "SE Suisun")) %>%
  mutate(Scenario = str_remove(Scenario, "_l95"))

zoopscon24 = left_join(zoopsmedian2024b, zoopsU952024b)%>%
                             left_join(zoopsl952024b) %>%
  mutate(Scenario = factor(Scenario, levels = c("NoAction", "X2Only", "GatesOnly", "Gates+X2")),
         Taxon = factor(IBMR, levels = c("pdiapfor", "pdiapjuv", "acartela", "eurytem", "euryjuv",
                                         "othcalad","othcaljuv", "limno", "othcyc", "allcopnaup", "daphnia", "othclad",
                                         "other"),
                        labels = c("Pseudodiaptomus", "Pseudodiaptomus juvenile", "Acartiella",
                                   "Eurytemora", "Eurytemora juvenile", "Other Calanoid Copepod", "Other Calanoid Juvenile",
                                   "Limnoithona", "Other Cylcopoid Copepod", "Copepod Nauplii",
                                   "Daphnia", "Other Cladocera", "Other")),
         Bigregion = case_match(region, c("SW Suisun", "SE Suisun", "NE Suisun") ~ "Suisun Bay",
                                                       c("NW Suisun") ~ "Grizzly bay",
                                                       "Suisun Marsh" ~ "Suisun Marsh",
                                                       "Confluence" ~ "Confluence")) %>%
   group_by(DOY, Taxon, Scenario, Bigregion) %>%
  summarize(Biomass = mean(Biomass, na.rm =T), BiomassU95 = mean(BiomassU95, na.rm =T),
            BiomassL95 = mean(BiomassL95, na.rm =T))

library(RColorBrewer)

mypal = c(brewer.pal(8, "Set1"),brewer.pal(8, "Dark2"))

zoopscon24b = mutate(zoopscon24, 
                     Month = case_when(DOY %in% c(153:182) ~ 6,
                       DOY %in% c(183:213) ~ 7,
                                      DOY %in% c(214:244) ~ 8,
                                      DOY %in% c(245:274) ~ 9,
                                      DOY %in% c(275:305) ~ 10)) %>%
  group_by(Month, Taxon, Scenario, Bigregion) %>%
  summarize(sdbio = sd(Biomass, na.rm =T), Biomass = median(Biomass, na.rm =T), upper = median(BiomassU95, na.rm =T),
            lower = median(BiomassL95, na.rm =T))

ggplot(filter(zoopscon24,Scenario != "Gates+X2"), aes(x = DOY, y = Biomass, fill = Taxon))+
  geom_area()+
  facet_grid(Bigregion~Scenario, scales = "free_y")+
  xlim(175, 305)+
  scale_fill_manual(values = mypal)+
  scale_x_continuous(breaks = c(183, 214, 245, 275), labels = c("Jul", "Aug", "Sep", "Oct"))+
  xlab("Day of Year")+ theme_bw()

ggplot(filter(zoopscon24b, Taxon == "Pseudodiaptomus", !is.na(Month)), aes(x = Bigregion, y = Biomass, fill = Scenario))+
  geom_col(position = "dodge")+ facet_wrap(~Month, scales = "free_y")+
  geom_errorbar(aes(group = Scenario, ymin = lower, ymax = upper), position = "dodge")

ggplot(filter(zoopscon24b, Taxon == "Pseudodiaptomus", !is.na(Month), Scenario != "Gates+X2"), aes(x = Month, y = Biomass, fill = Scenario))+
  geom_col(position = "dodge")+ facet_wrap(~Bigregion, scales = "free_y")+
  geom_errorbar(aes(group = Scenario, ymin = lower, ymax = upper), position = "dodge")+
  ylab("Median predicted \nbiomass of Pseudodiaptomus")

```
Why does growth go down with x2 in some of the suisun regions?


```{r}
zoopscon24 = left_join(zoopsmedian2024b, zoopsU952024b)%>%
                             left_join(zoopsl952024b) %>%
  mutate(Scenario = factor(Scenario, levels = c("NoAction", "X2Only", "GatesOnly", "Gates+X2")),
         Taxon = factor(IBMR, levels = c("pdiapfor", "pdiapjuv", "acartela", "eurytem", "euryjuv",
                                         "othcalad","othcaljuv", "limno", "othcyc", "allcopnaup", "daphnia", "othclad",
                                         "other"),
                        labels = c("Pseudodiaptomus", "Pseudodiaptomus juvenile", "Acartiella",
                                   "Eurytemora", "Eurytemora juvenile", "Other Calanoid Copepod", "Other Calanoid Juvenile",
                                   "Limnoithona", "Other Cylcopoid Copepod", "Copepod Nauplii",
                                   "Daphnia", "Other Cladocera", "Other")),
          Month = case_when(DOY %in% c(153:182) ~ 6,
                       DOY %in% c(183:213) ~ 7,
                                      DOY %in% c(214:244) ~ 8,
                                      DOY %in% c(245:274) ~ 9,
                                      DOY %in% c(275:305) ~ 10)) %>%
  group_by(Month, Taxon, Scenario, region) %>%
  summarize(sdbio = sd(Biomass, na.rm =T), Biomass = median(Biomass, na.rm =T), upper = median(BiomassU95, na.rm =T),
            lower = median(BiomassL95, na.rm =T))


ggplot(filter(zoopscon24, !is.na(Month)), aes(x = Scenario, y = Biomass, fill = Taxon)) + geom_col()+ 
  facet_grid(region~Month, scales = "free_y")

```

limnoithona

```{r}

limnoplot = filter(sal_conversions, IBMR == "limno", doy %in% c(153, 160, 170, 180, 190, 200, 210, 220, 230, 240, 250, 260, 270, 290, 304)) %>%
   mutate(across(starts_with("draw_"), ~exp(.x)-1))%>%
  pivot_longer(starts_with("draw_"), names_prefix="draw_", names_to="draw", values_to="fit")%>%
  mutate(fit=if_else(fit<0, 0, fit))%>%
  group_by(Region, Month, doy, SalSurf)%>%
  summarise(median= median(fit, na.rm=T), 
                        l95=quantile(fit, 0.025, na.rm=T), 
                        u95=quantile(fit, 0.975, na.rm=T))

  
ggplot(limnoplot, aes(x = SalSurf, y = median)) + geom_line()+
  facet_grid(doy~Region)
```

```{r}
limno2 = filter(zoop_data_mass, IBMR == "limno", Subregion_edit %in% c("NW Suisun", "East Suisun", "Suisun Marsh"))

ggplot(limno2, aes(x = SalSurf, y = BPUE_log1p)) + geom_smooth()+ facet_wrap(~Subregion_edit)
```

